---
title: "Exploring 2016 San Diego Police Stop Data"
output:
  html_document: default
  html_notebook: default
---

The purpose of this R Notebook is to clean and explore the 2016 police vehicle stop data that can be found at [Open San Diego's website](https://data.sandiego.gov/datasets/police-vehicle-stops/). 

```{r load packages, message = F}
library(tidyverse, verbose = T)
library(knitr, verbose = T)
library(stringr, verbose = T)
```

```{r set options, message = F, echo = F}
opts_chunk$set(message = F)
```

## About the Data 

There are four datasets that we'll be reading in, but two are for reference. The `race` dataframe describes the race codes used. 

```{r read race table}
race <- read_csv("http://seshat.datasd.org/pd/vehicle_stops_race_codes.csv")
```

The `dict` dataframe is a description of all the variables in the main police vehicle stop dataset. 

```{r read dictionary}
dict <- read_csv("http://seshat.datasd.org/pd/vehicle_stops_dictionary.csv")
```
 
This is the main vehicle stops dataset from 2016, which will require some cleaning. 

```{r read stops}
s_url <- "http://seshat.datasd.org/pd/vehicle_stops_2016_datasd.csv"

stops <- read_csv(s_url, col_types = cols(stop_id = col_character(),
                                          subject_age = col_character()), 
                  na = "NA", progress = F)
```

A lot of the character values will need to be cleaned up, including consolidating values that were spelled differently or capitalized. 

```{r}
# Yes and no's
stops$searched <- str_replace_all(stops$searched, 
                                  c("n" = "N", "\\\\" = "", "NA" = ""))
stops$arrested <- str_replace_all(stops$arrested, 
                                  c("n" = "N", "y" = "Y"))

# Remove "no age"
stops <- stops %>% 
  mutate(subject_age = ifelse(subject_age == "No Age", "", subject_age), 
         subject_age = as.numeric(subject_age))

# Remove X's in the gender field
stops <- stops %>% 
  mutate(subject_sex = ifelse(subject_sex == "" | subject_sex == "X", 
                              NA, 
                              subject_sex))

# Consolidate values stop_cause 
stops$stop_cause <- str_replace_all(stops$stop_cause, 
                        c("Bicycle Bicycle" = "Bicycle", 
                          "BICYCLE" = "bicycle", 
                          "Muni, County, H&S Code" = "MUNI, County, H&S Code"))
```

Let's examine the structure of the `stops` data. 

```{r}
str(stops)
```

There's also a dataframe that describes the actions taken during searches for each `stop_id`. 

```{r read details}
d_url <- "http://seshat.datasd.org/pd/vehicle_stops_search_details_2016_datasd.csv"

details <- read_csv(d_url, 
                    col_types = cols(stop_id = col_character(), 
                                     search_details_id = col_character(),
                                     search_details_type = col_character(),
                                     search_details_description = col_character()))
```

```{r echo = F, result = "hide"}
str(details)
```

Doing the `left_join` with `stops` and `details` results in multiple `stop_id` matches, probably because there are multiple observations in `details` for each of the `stop_id`. The result is that new rows are created in the resulting `comb` dataset.

```{r merge}
comb <- left_join(stops, details, by = "stop_id")
dim(stops); dim(details); dim(comb)
```

Let's focus first on analyzing the `stops` dataset without joining to the `details` dataset so that we don't misrepresent the number of observations originally collected in the `stops` dataset. 

## Exploratory Analysis

The following are exploratory plots to help generate hypotheses and ideas for feature selection. 

```{r}
ggplot(data = stops, aes(x = subject_sex)) + 
  geom_bar(fill = "cyan4") + 
  labs(title = "2016 San Diego Police Vehicle Stops by Gender", 
       subtitle = "data: Open San Diego", 
       x = "Gender", 
       y = "")
```

This boxplot drops rows because of NA values in the `subject_age` column. 

```{r}
ggplot(data = stops, aes(y = subject_age, x = subject_sex)) + 
  geom_boxplot(fill = "cyan4") + 
  scale_x_discrete("Gender", labels = c("M" = "Male", "F" = "Female", "NA")) +
  scale_y_continuous("Age") +
  labs(title = "2016 San Diego Police Vehicle Stops Median Age", 
       subtitle = "data: Open San Diego")
```

```{r}
ggplot(data = stops, aes(x = searched)) + 
  geom_bar(fill = "cyan4") + 
  scale_x_discrete("searched", labels = c("N" = "Not Searched", 
                                          "Y" = "Searched", "No data")) +
  labs(title = "2016 San Diego Police Vehicle Stops by Search Status", 
       subtitle = "data: Open San Diego", 
       x = "Was the subject searched?", 
       y = "")
```

When comparing the causes for the stops, we only considered the causes that had more than 10 instances. 

```{r}
count(stops, stop_cause) %>% 
  filter(n > 10) %>% 
  ggplot(aes(x = reorder(stop_cause, n), y = n)) + 
  geom_bar(stat = "identity", fill = "cyan4") +
  coord_flip() +
  labs(title = "2016 San Diego Police Vehicle Stops Count of Stop Cause",
       subtitle = "data: Open San Diego", 
       x = "", 
       y = "")
```

```{r}
ggplot(stops, aes(x = reorder(service_area, service_area, 
                              function(x) {length(x)}))) + 
  geom_bar(fill = "cyan4") +
  coord_flip() + 
  labs(title = "2016 San Diego Police Vehicle Stops by Service Area", 
       subtitle = "data: Open San Diego", 
       x = "Service Area Code", 
       y = "")
```