---
title: "Exploring 2016 San Diego Police Stop Data"
output:
  html_document: default
  html_notebook: default
---

The purpose of this R Notebook is to clean and explore the 2016 police vehicle stop data that can be found at [Open San Diego's website](https://data.sandiego.gov/datasets/police-vehicle-stops/). 

```{r load packages, message = F}
library(tidyverse, verbose = T)
library(knitr, verbose = T)
library(stringr, verbose = T)
```

```{r set options, message = F, echo = F}
opts_chunk$set(message = F)
```

## About the Data 

There are four datasets that we'll be reading in, but two are for reference. The `race` dataframe describes the race codes used. 

```{r read race table}
race <- read_csv("p_stop_data/vehicle_stops_race_codes.csv")
```

The `dict` dataframe is a description of all the variables in the main police vehicle stop dataset. 

```{r read dictionary}
dict <- read_csv("p_stop_data/vehicle_stops_dictionary.csv")
```
 
This is the main vehicle stops dataset, which will require some cleaning. 

```{r read stops}
stops <- read_csv("p_stop_data/vehicle_stops_2016_datasd.csv", 
                  col_types = cols(
                    stop_id = col_character(), 
                    subject_age = col_character()), 
                  na = "NA", 
                  progress = F)
```

Some of the string values in the `searched` and `arrested` fields need to be cleaned up so that the only values used are `N`, `Y`, or blank.

```{r}
stops$searched <- str_replace_all(stops$searched, c("n" = "N", "\\\\" = "", "NA" = ""))
stops$arrested <- str_replace_all(stops$arrested, c("n" = "N", "y" = "Y"))

unique(stops$searched); unique(stops$arrested)
```

Let's clean up the age variable by removing the "no age" values and converting it to a numeric class. 

```{r}
stops <- stops %>% 
  mutate(subject_age = ifelse(subject_age == "No Age", "", subject_age), 
         subject_age = as.numeric(subject_age))
```

Let's clean up the gender variable by converting blank fields and "X" values to NAs. 

```{r}
stops <- stops %>% 
  mutate(subject_sex = ifelse(subject_sex == "" | subject_sex == "X", 
                              NA, 
                              subject_sex))
```

Let's examine the structure of the `stops` data. 

```{r}
str(stops)
```

There's also a dataframe that describes the actions taken during searches for each `stop_id`. 

```{r read details}
details <- read_csv("p_stop_data/vehicle_stops_search_details_2016_datasd.csv",
                    col_types = cols(
                      stop_id = col_character(), 
                      search_details_id = col_character(), 
                      search_details_type = col_character(), 
                      search_details_description = col_character()
                    ))
```

```{r echo = F, result = "hide"}
str(details)
```

Doing the `left_join` with `stops` and `details` results in multiple `stop_id` matches, probably because there are multiple observations in `details` for each of the `stop_id`. The result is that new rows are created in the resulting `comb` dataset.

```{r merge}
comb <- left_join(stops, details, by = "stop_id")
dim(stops); dim(details); dim(comb)
```

Let's focus first on analyzing the `stops` dataset without joining to the `details` dataset so that we don't misrepresent the number of observations originally collected in the `stops` dataset. 

## Exploratory Analysis

```{r}
ggplot(data = stops) + 
  geom_bar(aes(x = subject_sex)) + 
  labs(title = "2016 San Diego Police Vehicle Stops by Gender", 
       subtitle = "data: Open San Diego", 
       x = "Gender", 
       y = "")
```

```{r}
ggplot(data = stops) + 
  geom_boxplot(aes(y = subject_age, x = subject_sex)) + 
  labs(title = "2016 San Diego Police Vehicle Stops Median Age", 
       subtitle = "data: Open San Diego", 
       x = "Gender", 
       y = "Age")
```

```{r}
ggplot(data = stops) + 
  geom_bar(aes(x = searched)) +
  labs(title = "2016 San Diego Police Vehicle Stops by Search Status", 
       subtitle = "data: Open San Diego", 
       x = "Was the subject searched?", 
       y = "")
```

```{r}
ggplot(data = stops, aes(
  x = reorder(stop_cause, stop_cause, function(x) {length(x)})
  )) + 
  geom_bar() +
  coord_flip() +
  labs(title = "2016 San Diego Police Vehicle Stops Count of Stop Cause",
       subtitle = "data: Open San Diego", 
       x = "", 
       y = "")
```

```{r}
ggplot(data = stops, aes(
  x = reorder(service_area, service_area, function(x) {length(x)})
)) + 
  geom_bar() +
  coord_flip() + 
  labs(title = "2016 San Diego Police Vehicle Stops by Service Area", 
       subtitle = "data: Open San Diego", 
       x = "Service Area Code", 
       y = ""
       )
```